Highly Available Transactions: Virtues and Limitations

### 2013-03-27 writeup/intro.tex
more intro text

Requirements for high availability, predictable low latency, and
simplicity of coordination protocols have led many recent distributed
database designers to renounce several guarantees provided by
traditional database systems~\cite{bigtable, pnuts, dynamo}. In
particular, many modern systems forgo providing ACID
guarantees---which often provide single-system image programmability
but suffer from unavailability and often expensive coordination
costs~\cite{davidson-survey}. The CAP Theorem~\cite{gilbert-cap},
demonstrating the incompatibility of ``always on'' operation with
guarantees on data item recency in the presence of unreliable
networks, has led a large number of systems to prioritize high
availability and guaranteed low latency, which are especially
important in geo-replicated contexts. We observe a dearth of highly
available systems providing general-purpose multi-operation
\textit{transactions} over multiple data items with useful semantics.

While serializability---the gold standard of traditional ACID
transactions---is not achievable with high availability for general
read-write transactions~\cite{davidson-survey}, databases have a long
tradition of providing weaker isolation and consistency
guarantees~\cite{adya, ansicritique, gray-isolation}. Due to their
concurrency and overall performance benefits, weak isolation models
are widely employed, overwhelmingly as the default setting in today's
traditional ``ACID'' and up-and-coming ``NewSQL'' databases and often
as the maximum guarantee offered (Section~\ref{sec:modernacid}). While
this weak isolation does not provide serializability for
general-purpose transactions, it is apparently strong enough to
deliver acceptable behavior to many application programmers and is
substantially stronger than the typical semantics of current highly
available systems. This raises the question: which of these semantics
can we provide while remaining highly available?

>>>>

Requirements for high availability, predictable low latency, and
simplicity of coordination protocols have led many recent distributed
database designers to renounce several guarantees provided by
traditional database systems~\cite{bigtable, pnuts, dynamo}. In
particular, many modern systems forgo providing ACID guarantees, which
often provide single-system image programmability but suffer from
unavailability and often expensive coordination
costs~\cite{davidson-survey}. In particular, the July 2000
announcement of the CAP Theorem~\cite{gilbert-cap}, which states that
a highly available system cannot provide recency guarantees in the
presence of network partitions, proved a rallying call for a
generation of scalable but semantically weak systems. Over a decade
later, we observe a dearth of highly available systems providing
general-purpose multi-operation \textit{transactions} over multiple
data items with useful semantics.

While serializability, the gold standard of traditional ACID
transactions, is not achievable with high availability for general
read-write transactions~\cite{davidson-survey}, databases have a long
tradition of providing weaker isolation and consistency
guarantees~\cite{adya, ansicritique, gray-isolation}. Due to their
concurrency and overall performance benefits, weak isolation models
are widely employed, overwhelmingly as the default setting in today's
traditional ``ACID'' and up-and-coming ``NewSQL'' databases and often
as the maximum guarantee offered (Section~\ref{sec:modernacid}). While
this weak isolation does not provide serializability for
general-purpose transactions, it is apparently strong enough to
deliver acceptable behavior to many application programmers and is
substantially stronger than the typical semantics of current highly
available systems. This raises the question: which of these semantics
can we provide while remaining highly available?
